%%
%% This is file `sample-sigconf.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `sigconf')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigconf.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%%
%% Commands for TeXCount
%TC:macro \cite [option:text,text]
%TC:macro \citep [option:text,text]
%TC:macro \citet [option:text,text]
%TC:envir table 0 1
%TC:envir table* 0 1
%TC:envir tabular [ignore] word
%TC:envir displaymath 0 word
%TC:envir math 0 word
%TC:envir comment 0 0
%%
%%
%% The first command in your LaTeX source must be the \documentclass
%% command.
%%
%% For submission and review of your manuscript please change the
%% command to \documentclass[manuscript, screen, review]{acmart}.
%%
%% When submitting camera ready or to TAPS, please change the command
%% to \documentclass[sigconf]{acmart} or whichever template is required
%% for your publication.
%%
%%
\documentclass[sigconf]{acmart}
%\documentclass[sigconf, nonacm]{acmart}


\copyrightyear{2024}
\acmYear{2024}
\setcopyright{rightsretained}
\acmConference[WWW '24]{Proceedings of the ACM Web Conference 2024}{May 13--17, 2024}{Singapore, Singapore}
\acmBooktitle{Proceedings of the ACM Web Conference 2024 (WWW '24), May 13--17, 2024, Singapore, Singapore}\acmDOI{10.1145/3589334.3645520}
\acmISBN{979-8-4007-0171-9/24/05}

\settopmatter{printacmref=true}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% For managing citations, it is recommended to use bibliography
%% files in BibTeX format.
%%
%% You can then either use BibTeX with the ACM-Reference-Format style,
%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
%% support for advanced citation of software artefact from the
%% biblatex-software package, also separately available on CTAN.
%%
%% Look at the sample-*-biblatex.tex files for templates showcasing
%% the biblatex styles.
%%

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

\usepackage{comment}
\usepackage{multirow}
\usepackage{amsmath,pifont}
\usepackage{booktabs}
\usepackage{rotating}
\usepackage{array}
\usepackage[para]{footmisc}
\usepackage{tablefootnote}
\usepackage{xcolor}
\usepackage{colortbl}
\usepackage{longtable}
\usepackage{circledsteps}
\usepackage{romannum}
\usepackage{longtable}
\usepackage{tcolorbox}
\usepackage{enumitem} % no indent [leftmargin=*]
\usepackage{balance}

\definecolor{Permissive}{HTML}{18AFCC}
\definecolor{Public}{HTML}{5F9099}
\definecolor{Copyleft}{HTML}{CC1865}

\hyphenation{OpenAI Model-Go} 
%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{ModelGo: A Practical Tool for Machine Learning License Analysis}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.

\author{Moming Duan}
\affiliation{%
  \institution{National University of Singapore}
  %\city{Singapore}
  \country{Singapore}}
\email{moming@nus.edu.sg}


\author{Qinbin Li}
\affiliation{%
  \institution{UC Berkeley}
  \city{Berkeley}
  \country{USA}}
\email{qinbin@berkeley.edu}

\author{Bingsheng He}
\affiliation{%
  \institution{National University of Singapore}
  %\city{Singapore}
  \country{Singapore}}
\email{hebs@comp.nus.edu.sg}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Moming Duan, Qinbin Li, \& Bingsheng He}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
Productionizing machine learning projects is inherently complex, involving a multitude of interconnected components that are assembled like LEGO blocks and evolve throughout development lifecycle.
These components encompass software, databases, and models, each subject to various licenses governing their reuse and redistribution.
However, existing license analysis approaches for Open Source Software (OSS) are not well-suited for this context.
For instance, some projects are licensed without explicitly granting sublicensing rights, or the granted rights can be revoked, potentially exposing their derivatives to legal risks.
Indeed, the analysis of licenses in machine learning projects grows significantly more intricate as it involves interactions among diverse types of licenses and licensed materials.
To the best of our knowledge, no prior research has delved into the exploration of license conflicts within this domain.
In this paper, we introduce ModelGo, a practical tool for auditing potential legal risks in machine learning projects to enhance compliance and fairness.
With ModelGo, we present license assessment reports based on five use cases with diverse model-reusing scenarios, rendered by real-world machine learning components.
Finally, we summarize the reasons behind license conflicts and provide guidelines for minimizing them.
Our code is publicly available at \url{https://github.com/Xtra-Computing/ModelGo}.
\end{abstract}

%bundling the LGPL-LR corpus dataset with another corpus dataset licensed under CC-BY-SA 4.0 can result in a copyleft proliferation conflict. %as stipulated by the derivative work terms of LGPL-LR.
%In contrast, bundling this corpus with a model licensed under copyleft GPL 3.0 does not lead to such conflict, as the derivative work is no longer considered a linguistic resource.

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
\begin{CCSXML}
  <ccs2012>
     <concept>
         <concept_id>10011007.10011074.10011134.10003559</concept_id>
         <concept_desc>Software and its engineering~Open source model</concept_desc>
         <concept_significance>300</concept_significance>
         </concept>
     <concept>
         <concept_id>10003456.10003457.10003580.10003585</concept_id>
         <concept_desc>Social and professional topics~Testing, certification and licensing</concept_desc>
         <concept_significance>500</concept_significance>
         </concept>
   </ccs2012>
\end{CCSXML}

\ccsdesc[500]{Social and professional topics~Testing, certification and licensing}
\ccsdesc[300]{Software and its engineering~Open source model}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{License analysis, AI licensing, model mining}

%\received{20 February 2007}
%\received[revised]{12 March 2009}
%\received[accepted]{5 June 2009}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\input{INTRODUCTION}

\input{BACKGROUND}

\input{METHOD}

\input{CASE}

\section{Conclusion}
\label{sec:conclusion}
Component reusing is prevalent in today's ML project development lifecycle, yet legal compliance issues are often ignored. Furthermore, it can be challenging for developers to understand elusive license terms and identify the potential risk of license violations.
Therefore, given the particularity of ML projects and licensing practices, we propose a practical license analysis tool to analyze their license conflicts.
We leverage five case studies to demonstrate the feasibility of our method, and our findings provide constructive guidelines to minimize conflicts.


%\section{Disclaimer}
%The content presented in this article is intended for general informational purposes only and should not be construed as legal advice. Any views, opinions, findings, conclusions, or recommendations expressed in this material are the sole responsibility of the author(s) and do not represent the perspectives of any organization or entity.


%%
%% The acknowledgments section is defined using the "acks" environment
%% (and NOT an unnumbered section). This ensures the proper
%% identification of the section in the article metadata, and the
%% consistent spelling of the heading.

\begin{acks}
This research is supported by the National Research Foundation Singapore and DSO National Laboratories under the AI Singapore Programme (AISG Award No: AISG2-RP-2020-018). 
Any opinions, findings and conclusions or recommendations expressed in this material are those of the authors and do not reflect the views of National Research Foundation, Singapore.
\end{acks}



%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
%\balance
\bibliography{REF}

%%
%% If your work has an appendix, this is the place to put it.
\newpage
\appendix
\input{APPENDIX}

\end{document}
\endinput
%%
%% End of file `sample-sigconf.tex'.
